---
title: "Reece_HW1"
author: "Tanner Reece"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(brms)
library(knitr)
library(rstatix)
library(effectsize)
library(tidybayes)
library(modelr)
library(formatR)
library(modeldb)
library(emmeans)
library(cowplot)
library(knitr)
library(cmdstanr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 100), tidy = TRUE)
```


```{r}
set.seed(74)
dat <- tibble(
  a = rep(0:1,each=150),
  b = as.factor(rep(rep(0:2,each=1),100)))

dat <- dat %>% 
  add_dummy_variables(b, values = c(0,1,2), remove_original=F)

y = 2 + rnorm(300,.7,.5)*dat$a + rnorm(300,.9,.5)*dat$b_1 +  rnorm(300,-.8,.5)*dat$b_2 + rnorm(300,.1,.5)*dat$a*dat$b_1 + rnorm(300,.19,.5)*dat$a*dat$b_2 + rnorm(300,0,.5)

dat <- dat %>% add_column(y)
```

# 1) Run a full factorial interaction model using brms with the above'dat' dataset.
# Interpret the summary output.
```{r}
# dat$a <- dat$a %>% dplyr::recode("0" = "1","1" = "2")
# dat$b <- dat$b %>% dplyr::recode("0" = "3")
dat$a <- dat$a %>% as.factor()

mod1 <- brm(family = gaussian,
             y ~ 0 + a:b,
             prior = c(prior(normal(0,5), class = b),
                       prior(cauchy(0,10), class = sigma)),
             data = dat,
             iter = 1000, warmup = 500, chains = 2, cores = 2,
            backend="cmdstanr")


summary(mod1)
prior_summary(mod1)
#Coefficient Interpretation
#The coefficients represent the cell means for each a/b combination and the sigma_ variables represent the estimated residual variances for each cell mean
```

# 2) Graph

## 2a) The two simple main effects
```{r}
plot_grid(
  emmeans(mod1, ~a) %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(x = a, y = .value))+
  stat_eye(fill="blue") +
  stat_summary(aes(group = NA), fun.y = mean, geom="line"),

emmeans(mod1, ~b) %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(x = b, y = .value))+
  stat_eye(fill="red") +
  stat_summary(aes(group = NA), fun.y = mean, geom="line"),

nrow=1,ncol=2
)

```

## 2b) A figure to demonstrate the interaction effects
```{r}
emmeans(mod1, ~ a:b) %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(x = a, y = .value,fill=b))+
  stat_eye()+
  stat_summary(aes(group = b), fun.y = mean, geom="line")
```

## 2c) The differences/contrasts between the simple main effects
```{r}
mod1 %>% 
emmeans(~a) %>% 
  contrast(method = "pairwise") %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(y = contrast, x = .value))+
  stat_halfeye(fill="blue", alpha = 0.5)

mod1 %>% 
emmeans(~b) %>% 
  contrast(method = "pairwise") %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(y = contrast,x=.value))+
  stat_halfeye(aes(fill=contrast), alpha = 0.5)+
  theme(legend.position="NULL")

```

## 2d) The differences/contrasts betwen each cell
```{r}
#All possible cell pairs
mod1 %>% 
emmeans(~a:b) %>% contrast(method="pairwise") %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  stat_halfeye(aes(fill=contrast), alpha = 0.5)+
  theme(legend.position="NULL")

#We can also look at the simple effects
#a1-a0 @ each level of b
mod1 %>% 
emmeans(~a|b) %>% contrast(method="pairwise",simple="each") %>% 
  gather_emmeans_draws() %>% 
  filter(.grid == 'simple contrasts for a') %>% 
  ggplot(aes(x = -1*.value, y = b, fill = b)) +
  stat_halfeye(aes(fill=b), alpha = 0.5)+
  xlab('a1-a0')+
  ggtitle('a1 - a0 at each level of b')+
  theme(legend.position="NULL")

#and b1 - b0, b2 - b0, and b2 - b1 @ each level of a
mod1 %>% 
emmeans(~a|b) %>% contrast(method="pairwise",simple="each") %>% 
  gather_emmeans_draws() %>% 
  filter(.grid == 'simple contrasts for b') %>% 
  dplyr::select(-b,-.chain,-.iteration) %>% 
  ggplot(aes(x = .value, y = contrast, fill = contrast)) +
  facet_wrap(~a) +
  stat_halfeye(aes(fill=contrast), alpha = 0.5)+
  theme(legend.position="NULL")
```

## 2f) A plot (your choice) but with standardized, not raw values

```{r}
# mod1 %>% 
#   spread_draws(b_a0:b0,b_a1:b0,sigma) %>% 
#   select(b_a0,b_a1,sigma) %>% 
#   rename(a0 = "b_a0", a1 = "b_a1") %>% 
#   mutate(std.effect = (a1-a0)/sigma) %>% 
#   ggplot(aes(x = std.effect))+
#   ylab(NULL)+
#   ggtitle("Cohen's d for a1 - a0 contrast")+
#   stat_dotsinterval()+
#   theme(axis.text.y=element_blank())

draws <- mod1 %>% 
  tidy_draws() 

draws$a0 <- rowMeans(draws[,c(4,6,8)], na.rm=TRUE)
draws$a1 <- rowMeans(draws[,c(5,7,9)], na.rm=TRUE)
draws %>% mutate(std.effect = (a1-a0)/sigma) %>% 
  ggplot(aes(x = std.effect))+
  ylab(NULL)+
  ggtitle("Cohen's d for a1 - a0 contrast")+
  stat_dotsinterval()+
  theme(axis.text.y=element_blank())

```

# 3) Test. Use a 97% CI

## 3a) What are the medians and the 97% CIs for the simple main effect estimates

```{r}
mod1 %>% 
emmeans(~a) %>% 
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)

#Median and 97% CIs:
#a0 = 2.02 (1.88,2.16)
#a1 = 2.95 (2.82, 3.09)

mod1 %>% 
emmeans(~b) %>% 
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97) 

#Median and 97% CIs:
#b0 = 2.36 (2.18,2.52)
#b1 = 3.37 (3.20, 3.55)
#b2 = 1.71 (1.54, 1.90)

```

## 3b) What are the medians and the 97% CIs for each of the cell estimates?
```{r}
mod1 %>% 
emmeans(~a:b) %>% 
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)

#Median and 97% CIs:
#a0b0 = 2.15 (1.89,2.39)
#a0b1 = 2.76 (2.51,3.03)
#a0b2 = 1.15 (0.91,1.39)
#a1b0 = 2.57 (2.36,2.80)
#a1b1 = 3.99 (3.75,4.26)
#a1b2 = 2.28 (2.03,2.51)

```

## 3c) Are any of the simple effects significantly different from one another

```{r}
#Main effect of a is significant. a0 - a1 -> -0.93 (-1.12,-0.72)
mod1 %>% 
emmeans(~a) %>% contrast(method="pairwise") %>%
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)




mod1 %>% 
emmeans(~b) %>% contrast(method="pairwise") %>%
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)
#b0-b1 -> -1.02 (-1.3,-0.78)
#b0-b2 -> 0.64 (0.39, 0.88)
#b1-b2 -> 1.66 (1.41,1.93)
#All pairwise comparisons within b are significant
                           
```

## 3d) Are any of the cell comparisons significantly different from one another?
```{r}
#We can look at all possible contrasts

mod1 %>% 
emmeans(~a:b) %>% contrast(method="pairwise") %>%
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)


#Or simple contrasts
#All simple effects are significant EXCEPT b0 - b2 @ a = 1
mod1 %>% 
emmeans(~a:b) %>% contrast(method="pairwise",simple="each") %>%
  gather_emmeans_draws() %>% 
  median_qi(.width = 0.97)
                           
```
# 4) Interpet the R2
```{r}
bayes_R2(mod1,summary=F) %>% as.data.frame() %>% 
  ggplot(aes(x = R2))+
  stat_dotsinterval()

bayes_R2(mod1,summary=T)
#~52.4% of the variance in y is explained by our predictors
```